Lucy.js: Client-Side Indexes for Fast Full-Text Searching

### 2014-12-11 paper/paper.tex
results section

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.4]{hybrid-trie}
  \caption{Structure of a hybrid trie with leaf and path shrink.}
\end{figure}

We also implemented a novel approach, which we call a hybrid trie: A hybrid between an inverted index and a prefix (or suffix) trie. This approach takes advantage of the fact that every object store in IndexedDB is essentially a hash index with very fast key lookups.

Every node in our hybrid trie has its entire path as its key. Therefore, when faced with a prefix query, we can directly jump to the position in the trie that corresponds to that prefix instead of doing \(O(n)\) lookups on the size of the prefix.

Then, every node includes an array of characters that are the children of the node in the trie. The key of each child is determined by concatenating the key of the parent node with each of the strings in the children array.

Every node can also potentially include a list of document ids, for documents that contain its prefix as a token (i.e. not a prefix of some longer token). Nodes without children must contain a list of document ids, but internal nodes can also contain document ids. For example, the node for “cat” would contain document ids of documents with the word "cat" itself, as well as children, for terms like "catsitting", "caturday" or "catnip". For these examples, the list of children could be \texttt{['n', 's', 'u']}.

\subsubsection{Leaf and path Shrink}

To eliminate wasteful lookups, we implement leaf and path shrinkage. In tries, there are often long sequences of nodes with only one child. If those nodes have no document ids, we can combine them to a single node without any loss of functionality.

Shrinkage is a post-processing step after index generation, since we cannot know which nodes can be combined until all documents have been added to the index. Then our shrinkage algorithm works as follows:

\begin{verbatim}
for each node: 
    if node.children = 1:
        for each parent's children:
        	    if children[i] points to node:
        	    	children[i] += child.prefix;
        // number of characters to reach parent:
        child.parent++
        delete node
\end{verbatim}

This results in a small speed increase, but since lookups by key in indexedDB are already quite fast, the increase is not significant. However, it also results in space savings, which is important because space on the client-side is very limited: after only a few Megabytes, browsers will ask for confirmation by the user, which is disruptive and can lead to non-technical users declining merely out of confusion.

@@

\begin{figure}
   \includegraphics[scale=.53]{query-times}
  \caption{Comparison of query times for various queries with different result set sizes (left to right: smaller to larger) between our three main types of indexes and sequential search (the only way to search strings that indexedDB offers). }
\end{figure}

\subsection{Depth-Limiting for Tries}
As seen in Figure~\ref{comptrie}, the maximum depth parameter has an effect on lookup times in trie indexes. The graph shows the results for a trie index prefix search, but similar results would be expected on suffix searches. A small set of input tweet data was used (367 KB data, about 2100 tweets), and the same three words were looked up in indexes created with differing maximum depths. The lookup time tends to decrease as the maximum depth increases; this is because depth cutoffs can result in slow searches through word mappings stored on leaf nodes (see Section~\ref{depthlimit} for details). The difference in times is more pronounced for the term ``feel'', with a drop from 26.78ms to 18.05ms. For the term ``james'', lookup times are much closer together and not as affected by depth limits. This is due to the fact that ``feel'' yielded many more results (85 results) than ``james'' (4 results), suggesting that more potential matches had to be searched at the leaf nodes for ``feel'' than for ``james''. Results for ``wait'' are between the others, and also show a moderate drop in lookup time as the maximum depth increases. This is most likely because ``wait'' produced a moderate 36 results, and therefore a moderate number of terms were searched at the leaves.

@@

\section{Evaluation and Applications}

\subsection{Client-side vs. Server-side Search}

\begin{figure}[h!]
   \includegraphics[scale=.53]{demo_results}
  \caption{caption here.}
\end{figure}

@@

\section{Future Work}
The functionality of Lucy.js can be expanded to perform more complex queries and to improve the quality of returned results. While prefix and suffix searches are currently handled, arbitrary substring or regex searches are not supported. With these new features, users could write more customized queries and get better results. Additional forms of natural language processing could also improve the quality of results returned on searches. For example, synonym replacement could be implemented similarly to Postgres , allowing Lucy.js to find text results with similar themes or meaning. To improve performance, the existing indexes -- particularly the trie index -- could be optimized during index creation.

Another avenue for future research would be collaborative searches between a client and server. A client could store partial data and use Lucy.js to search locally, while the server could respond with additional results not stored on the client. This could be useful to perform client-side caching of data for a web application, so that communication with the server is not always required.

>>>>

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.4]{hybrid-trie}
  \caption{Structure of a hybrid trie.}
\end{figure}

@@

\begin{figure}[h!]
   \includegraphics[scale=.53]{query-times}
  \caption{caption here}
\end{figure}

\subsection{Depth-Limiting for Tries}
As seen in Figure~\ref{comptrie}, the maximum depth parameter has an effect on lookup times in trie indexes. The graph shows the results for a trie index prefix search, but the similar results would be expected on suffix searches. A small set of input tweet data was used (367 KB data, about 2100 tweets), and the same three words were looked up in indexes created with differing maximum depths. The lookup time tends to decrease as the maximum depth increases; this is because depth cutoffs can result in slow searches through word mappings stored on leaf nodes (see Section~\ref{depthlimit} for details). The difference in times is more pronounced for the term ``feel'', with a drop from 26.78ms to 18.05ms. For the term ``james'', lookup times are much closer together and not as affected by depth limits. This is due to the fact that ``feel'' yielded many more results (85 results) than ``james'' (4 results), suggesting that more potential matches had to be searched at the leaf nodes for ``feel'' than for ``james''. Results for ``wait'' are between the others, and also show a moderate drop in lookup time as the maximum depth increases. This is most likely because ``wait'' produced a moderate 36 results, and therefore a moderate number of terms were searched at the leaves.

@@

\section{Evaluation}

As seen previously, our implemented indexes perform much faster than a naive sequential scan over all full-text values. Given the current state of the IndexedDB API, this constitutes a great improvement in new functionality for fully client-side applications. However, most web applications are not client-side only and have a server component that can then query a server-side search engine such as Lucene, MySQL, or PostgreSQL. Thus in our evaluation, we seek to determine whether the client-side approach is faster than the server-side approach. While Lucene may generally be faster than Lucy.js for query evaluation, which is likely given that it has special purpose data structures and many years of development, a fully client-side approach may still beat a server-side approach due to network latency.

The server we built is implemented in Python and uses Django, Apache, and mod\_wsgi. It is hooked up to a MySQL database which has a full-text index on the tweet text. Both are hosted on the OpenStack cloud in a virtual machine, with the database residing in the virtual machine's local storage. The full-text search bar on the client side uses our inverted index to search for the query, while the typeahead search bar uses our prefix trie as it is more suited to finding prefixes.

We conduct our evaluation on two different real-life applications: a full-text search bar and a typeahead search bar that makes queries for every character as they are typed. The dataset that we test on is the same Twitter dataset used previously. We test each application on two different size datasets of 100 and 2000 tweets picked randomly. For the search queries, because we do not have access to typical search queries in real life, we simply take a random sample of 20 non-stopword words from the full tweet dataset for each set of experiments. We conduct 8 trials for each condition and delete and reinsert and recreate indexes before each trial. For the full-text search bar, we simply search for the 20 words and record the time taken. For the typeahead bar, we search for each of the 20 words as well as all prefixes of each word, in order as if one were typing into the search bar, and record the time.

\begin{figure}[h!]
   \includegraphics[scale=.43]{figure_1}
  \caption{Comparison of client-side versus server-side indexes for full-text search. Tested on a full-text search bar (FT Search) and a typeahead search bar (Typeahead) and with Twitter dataset sizes of 100 and 2000 tweets.}
\label{compare}
\end{figure}

\subsection{Results}

In Figure~\ref{compare}, we report the average time taken for each of the approaches (client versus server) on the two applications and dataset sizes. As can be seen, overall the client-side only approach is faster than than the server-side approach. When looking at full-text search, even when the dataset size grows over a magnitude larger, the server-side average time varies little. This suggests that the time taken is mostly taken up by network latency. We do see a small increase in the time taken for the client-side database. This suggests that as datasets grow larger, it may be a better decision to store the majority of the data on the server side and do queries on smaller subsets in the client.

When looking at the typeahead search bar, we see that when the dataset gets larger, the average time taken for both client and server increases, though the client-side increase is greater. When looking more closely at the times for different queries made, we notice that there is a great deal of variation in the time because a very short prefix will return a much larger result set than a longer prefix. The client side had greater variation overall, with a standard deviation of 56.477ms, maximum time of 256ms, and minimum time of 1ms. The server side had a standard deviation of 49.116ms, maximum time of 220ms, and minimum time of 14ms. Given these findings, a best approach may be employ a mixed approach and send shorter prefixes to the server and longer prefixes to the client. Generally, queries that will return a small result set would be faster if they are served from the client, as the network latency always tacks on extra time for the query to the server. Queries that will return a very large result set may have better performance if they are sent to the server.

\subsection{Live Demos}

\begin{figure}[h!]
\centering
   \includegraphics[scale=.34]{typeahead}
  \caption{Screenshot of typeahead search bar demo.}
\label{demo}
\end{figure}
Demonstrations of the two applications are accessible at \texttt{http://128.52.184.3/}. Figure~\ref{demo} has a screenshot of the typeahead search bar demo. For both demos, the user first deletes all data on the server as well as on the client. Then they choose whether to import the data, existing as a JSON file on the server, into the server-side MySQL database or into the client-side IndexedDB database. At that point the typeahead bar will work, querying the respective database, returning results, and outputting the time taken for each query. The user can additionally do a time test, which runs a selection of queries against the respective database and outputs the average time. Since IndexedDB does not yet support a synchronous API, the time test for the client-side database requires the user to continue clicking the button to run more queries and see an updating average time.

@@

\section{Future Work}
